---
title: "data_manipulation"
author: "Helen Guan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Import FAS csv files

Import my first csv files.

```{r}
litters_data = read_csv("./datasets/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./datasets/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

## Select variables

```{r}
select(litters_data, group, litter_number, gd0_weight)

select(litters_data, litter_number, gd0_weight, pups_survive)
```

... by specifying a range:

```{r}
select(litters_data, gd_of_birth:pups_survive)
select(litters_data, group, gd_of_birth:pups_survive)
```

... by saying what you want to remove:

```{r}
select(litters_data, -litter_number)
```

```{r}
select(litters_data, group, litter_number, gest_day_0_weight = gd0_weight)
```

... use select helpers:
```{r}
select(litters_data, starts_with("gd"))

select(litters_data, litter_number, starts_with("gd"))

select(litters_data, litter_number, everything())
```

learning assessment
```{r}
select(pups_data, litter_number, sex, pd_ears)
```

## Filter observations

Do some filtering!!
```{r}
filter(litters_data, gd0_weight < 25)
filter(litters_data, gd0_weight >= 25)

filter(litters_data, pups_born_alive == 8)

filter(litters_data, !is.na(gd0_weight))

filter(litters_data, group == "Low8" | group == "Low7")
filter(litters_data, group %in% c("Low7", "Low8"))

filter(litters_data, pups_born_alive %in% 4:6)
```

learning assessment
```{r}
filter(pups_data, sex == 1)
filter(pups_data, sex == 2, pd_walk < 11)
```

## Mutate

Create variables

```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       group = tolower(group)
)

mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       wt_gain_squared = wt_gain^2
)
```

## Arrange

Arrange the data...

```{r}
arrange(litters_data, gd0_weight)

arrange(litters_data, desc(pups_born_alive), gd0_weight)
```

## Piping

First option:

```{r}
litters_data_raw = read_csv("./datasets/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = mutate(litters_data_selected_cols, 
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group))
litters_data_with_vars
```

Used nested function calls:
```{r}
litters_data_clean = 
  mutate(
    select(
      janitor::clean_names(
        read_csv("./datasets/FAS_litters.csv", col_types = "ccddiiii")
        ), 
    -pups_survive
    ),
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
  )
litters_data_clean
```

Use piping
```{r}
read_csv("./datasets/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
  ) %>%
  filter(.data = ., !is.na(gd0_weight))
```

Illustrate placeholder use with "lm"
```{r}
read_csv("./datasets/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
  ) %>%
  filter(.data = ., !is.na(gd0_weight))
```



